### 2009-08-07 script/en/script-en.tex
fixed overfull hbox

% $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,11pt,oneside]{book}
\overfullrule=5pt % uncomment to mark overfull boxes
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{pdfpages}   % SR_2009: added new cover

\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
%\bibliographyunit[\section] % bibunits für das ganze Dokument aktivieren
\defaultbibliography{references}
\defaultbibliographystyle{abbrv}
\usepackage{float}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,hypertexnames=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels and hypertexnames=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{ruled}
\newfloat{sagecode}{!h}{loc}[chapter]
\floatname{sagecode}{Sage sample}
\newfloat{cryptoprocedure}{!h}{lop}[chapter]
\floatname{cryptoprocedure}{Crypto Procedure}

\makeindex
\hypersetup{colorlinks=true}

% \setlength{\textwidth}{460 pt}
% \hoffset-0.7in
\newtheorem{definition}{Definition}[section]
%\newcommand{\href}[2]{#2}
\newcommand{\hypertag}[2]{#1}
%\newcommand{\hyperlink}[2]{#2}
%\newcommand{\hypertarget}[2]{#2}
\newtheorem{theorem}{Theorem}[section]
\newenvironment{Proof}[1]{\noindent\textbf{Proof #1} \\}{\hfill$\Box$\par}
\newenvironment{example}[1]{\noindent\textbf{Example#1}}{}
\newenvironment{remark}[1]{\noindent\textbf{Comment#1}}{}

% Define title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{ % Keine Leerzeilen, sonst Meldung: "Paragraph ended before \title was complete."
	{\Huge The CrypTool Script: \\
	 \vspace{5pt}
	Cryptography, Mathematics and More} \\
        \vspace{20pt}
	{\LARGE	Background reading\\ 
        for the free e-learning program CrypTool} \\
        \vspace{15pt}
        {\large (10th edition -- distributed with CrypTool version 1.4.30)}
        \vspace{180pt}
}

% Hier wird \vspace{32pt} einfach ignoriert
\author{
Copyright (c) Prof.\ Bernhard Esslinger (co-author and editor) \\
and the CrypTool Development Team, 1998-2009 \\
Frankfurt am Main, Germany\\*[1ex]
\url{www.cryptool.org}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\includepdf[page=1]{figures/coverv1_en.pdf}    % SR_2009: added new cover

\frontmatter

\pagestyle{plain}
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}
\addtolength{\footskip}{8pt}
%\setlength{\footskip}{4cm}
%\renewcommand{\footnoterule}{\parindent0cm\rule{13cm}{.1pt}\vspace{.2cm}}
%Formatierung der Fu"snoten

%space between text and footnote 
\renewcommand\footnoterule{%
  \vspace{2em}%   <-- one line space between text and footnoterule
%\kern-3\p@
  \hrule width .4\columnwidth
 \vspace{4pt}
%kern 2.6\p@
}

%\long\def\@makefntext#1{%
%    \parindent 1em%
%    \noindent
%    \hbox to 1.8em{\hss\@makefnmark}#1}

\maketitle

% Vergleichbar zur Nutzung von Texten aus der Wikipedia (=GNU Free Documentation Licence, GFDL)
% \vspace{200pt}
\newpage
\noindent This is a free document, so the content of the document can
be copied and distributed, also for commercial purposes --- as long as
the authors, title and the CrypTool web site (\url{www.cryptool.org})
are acknowledged. Naturally, citations from the CrypTool script are
possible, as in all other documents.\\
The specific license for this document is the GNU Free Documentation Licence
\hyperlink{appendix-GNU-fdl}{(see appendix)}.
% (see \ref{s:appendix-GNU-fdl}).

\begin{quote}
    Copyright \copyright{} 1998--2009 Bernhard Esslinger and the
    CrypTool Development Team. Permission is granted to copy,
    distribute and/or modify this document under the terms of the GNU
    Free Documentation License, Version 1.3 or any later version
    published by the Free Software Foundation; with no Invariant
    Sections, no Front-Cover Texts, and no Back-Cover Texts. A copy of
    the license is included in the section entitled ``GNU Free
    Documentation License''.
\end{quote}

\newpage
\input{chapters/aboutcryptool.tex}

\newpage
\pdfbookmark[0]{\contentsname}{Contents}
\tableofcontents

\newpage
\normalsize
%\parindent 0cm
\parskip 4pt

\input{chapters/introduction.tex}

\mainmatter

\input{chapters/cryptomethods.tex}
\input{chapters/paper_and_pencil.tex}
\input{chapters/primes.tex}
\input{chapters/numbertheory.tex}
\input{chapters/moderncryptography.tex}
\input{chapters/digitalsignatures.tex}
\input{chapters/ellipticcurves.tex}
\bibliographyunit[\chapter]
\bibliography*{../de/references}
\bibliographystyle*{plain}
\input{chapters/crypto2020.tex}
\bibliographyunit

\begin{appendix}
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{appendix-start}{}\label{s:appendix-start}
\chapter{Appendix}
    \begin{itemize}
      \item[1] \hyperlink{appendix-menutree}{CrypTool Menu Tree}
      \item[2] \hyperlink{appendix-authors}{Authors of the CrypTool Script}
      \item[3] \hyperlink{appendix-authors}{Bibliography of Movies and
                     Fictional Literature with Relation to Cryptography,
                     Books for Kids with Collections of Simple Ciphers}
      \item[4] \hyperlink{appendix-Learn-NT}
                     {Learning Tool for Elementary Number Theory}
      \item[5] \hyperlink{appendix-using-sage}
                     {Using Sage with this Script}
    \end{itemize}
  \newpage
  \input{chapters/menus.tex}
  \input{chapters/authors.tex}
  \input{chapters/what-is-sage.tex}
\end{appendix}

\input{chapters/gnu-fdl-1.3.tex}
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begingroup
% remove extra vertical space from \listoffigure and \listoftables so that 
% it look similar to \listof{sagecode}
%\setlength{\parskip}{0pt}
\renewcommand*{\addvspace}[1]{}

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{\listfigurename}
\listoffigures

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{\listtablename}
\listoftables

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Crypto Procedures}
\listof{cryptoprocedure}{List of Crypto Procedures}
\endgroup

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Sage Code Examples}
\listof{sagecode}{List of Sage Code Examples}

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Index}
\printindex

\end{document}