### 2012-08-20 script/en/script-en.tex

% $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,11pt,oneside]{book}
\overfullrule=5pt % uncomment to mark overfull boxes
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{pdfpages}   % SR_2009: added new cover

\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
%\bibliographyunit[\section] % bibunits für das ganze Dokument aktivieren
\defaultbibliography{references}
\defaultbibliographystyle{abbrv}
\usepackage{float}
\usepackage{subfig}
\usepackage{shorttoc}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{ruled}
\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{Sage sample}
\newfloat{cryptoprocedure}{!ht}{lop}[chapter]
\floatname{cryptoprocedure}{Crypto Procedure}

\makeindex
\hypersetup{colorlinks=true}

% \setlength{\textwidth}{460 pt}
% \hoffset-0.7in
\newtheorem{definition}{Definition}[section]
%\newcommand{\href}[2]{#2}
\newcommand{\hypertag}[2]{#1}
%\newcommand{\hyperlink}[2]{#2}
%\newcommand{\hypertarget}[2]{#2}
\newtheorem{theorem}{Theorem}[section]
\newenvironment{Proof}[1]{\noindent\textbf{Proof #1} \\}{\hfill$\Box$\par}
\newenvironment{example}[1]{\noindent\textbf{Example#1}}{}
\newenvironment{remark}[1]{\noindent\textbf{Comment#1}}{}

% Define title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{ % Keine Leerzeilen, sonst Meldung: "Paragraph ended before \title was complete."
	{\Huge The CrypTool Script: \\
	 \vspace{5pt}
	 Cryptography, Mathematics, and More} \\
        \vspace{20pt}
	{\LARGE	Background reading for CrypTool\\ 
         the free e-learning program\\
	 (with number theory code samples for Sage)} \\
        \vspace{15pt}
        {\large (11th edition -- distributed with CrypTool version 1.4.31)}
        \vspace{180pt}
}

% Hier wird \vspace{32pt} einfach ignoriert
\author{ % Die Klammer zur Fontänderung funktioniert bei author nur einzeilig im Gegensatz zu title?!
        {\LARGE Copyright (c) Prof.\ Bernhard Esslinger (co-author and editor)}\\[5pt]
        {\LARGE and the CrypTool Development Team, 1998-2012}\\[5pt]
        {\LARGE Frankfurt am Main, Germany} \\*[15pt]
        \url{www.cryptool.org}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\VerbatimFootnotes

% \includepdf[page=1]{figures/coverv1_en-300dpi.pdf}
% \includepdf[page=1]{figures/coverv1_en.pdf}
\includepdf[page=1]{figures/coverv1_en_draft.pdf}

\frontmatter

\pagestyle{plain}
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}
\addtolength{\footskip}{8pt}
%\setlength{\footskip}{4cm}
%\renewcommand{\footnoterule}{\parindent0cm\rule{13cm}{.1pt}\vspace{.2cm}}
%Formatierung der Fu"snoten

%space between text and footnote 
\renewcommand\footnoterule{%
  \vspace{2em}%   <-- one line space between text and footnoterule
%\kern-3\p@
  \hrule width .4\columnwidth
 \vspace{4pt}
%kern 2.6\p@
}

%\long\def\@makefntext#1{%
%    \parindent 1em%
%    \noindent
%    \hbox to 1.8em{\hss\@makefnmark}#1}

\maketitle

% Vergleichbar zur Nutzung von Texten aus der Wikipedia (=GNU Free Documentation Licence, GFDL)
% \vspace{200pt}
\begin{quote}
This is a free document, so the content of the document can
be copied and distributed, also for commercial purposes --- as long as
the authors, title and the CrypTool web site (\url{www.cryptool.org})
are acknowledged. Naturally, citations from the CrypTool script are
possible, as in all other documents.\\
The specific license for this document is the GNU Free Documentation Licence.

    Copyright \copyright{} 1998--2012 Bernhard Esslinger and the
    CrypTool Development Team. Permission is granted to copy,
    distribute and/or modify this document under the terms of the GNU
    Free Documentation License, Version 1.3 or any later version
    published by the Free Software Foundation; with no Invariant
    Sections, no Front-Cover Texts, and no Back-Cover Texts. A copy of
    the license is included in the section entitled
    \hyperlink{appendix-GNU-fdl}{``GNU Free Documentation License''}.
\end{quote}

\vspace{460pt}
Source cover photograph: \url{www.photocase.com}, Andre Guenther\\

Typesetting software: \LaTeX

Version control software: Subversion

\input{chapters/aboutcryptool.tex}

\newpage
\pdfbookmark[0]{Contents Overview}{ShortContents}
\shorttoc{Contents Overview}{0}
\clearpage\phantomsection
\pdfbookmark[0]{\contentsname}{Contents}
\tableofcontents

\newpage
\normalsize

%\parindent 0cm
% \parskip 4pt
% Setting the default for the vertical distance before a paragraph as distinguished variable
\newcounter{mycounterDefaultParskip}
\setcounter{mycounterDefaultParskip}{4}
\parskip \value{mycounterDefaultParskip} pt

\input{chapters/introduction.tex}

\mainmatter

\input{chapters/cryptomethods.tex}
\input{chapters/paper_and_pencil.tex}
\input{chapters/primes.tex}
\input{chapters/numbertheory.tex}
\input{chapters/moderncryptography.tex}
\input{chapters/digitalsignatures.tex}
\input{chapters/ellipticcurves.tex}
\bibliographyunit[\chapter]
\bibliography*{../de/references}
\bibliographystyle*{plain}
\input{chapters/crypto2020.tex}
\bibliographyunit

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begin{appendix}
\newpage
\hypertarget{appendix-start}{}\label{s:appendix-start}
\chapter{Appendix}
    \begin{itemize}
      \item[1] \hyperlink{appendix-menu-overview-CT1}{CrypTool 1 Menu Tree}
      \item[2] \hyperlink{appendix-template-overview-CT2}{CrypTool 2 Templates}
      \item[3] \hyperlink{appendix-function-overview-JCT}{JCrypTool Functions}
      \item[4] \hyperlink{appendix-authors}{Bibliography of Movies and
                     Fictional Literature with Relation to Cryptography,
                     Books for Kids with Collections of Simple Ciphers}
      \item[5] \hyperlink{appendix-Learn-NT}
                     {Learning Tool for Elementary Number Theory}
      \item[6] \hyperlink{appendix-using-sage}
                     {Using Sage with this Script}
      \item[7] \hyperlink{appendix-authors}{Authors of the CrypTool Script}
    \end{itemize}
  \newpage
  \input{chapters/menus.tex}
  \input{chapters/movies.tex}
  \input{chapters/what-is-sage.tex}
  \input{chapters/authors.tex}
\end{appendix}

\input{chapters/gnu-fdl-1.3.tex}

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begingroup
% remove extra vertical space from \listoffigure and \listoftables so that 
% it look similar to \listof{sagecode}
%\setlength{\parskip}{0pt}
\renewcommand*{\addvspace}[1]{}

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{\listfigurename}
\listoffigures

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{\listtablename}
\listoftables

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Crypto Procedures with Pseudo Code}
\listof{cryptoprocedure}{List of Crypto Procedures with Pseudo Code}
\endgroup

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Sage Code Examples}
\listof{sagecode}{List of Sage Code Examples}
\label{sc:List-of-Sage-Code-Examples}

\vspace{30pt}
\begin{itemize}
  \item All samples have been tested with Sage v 4.7 (release date 2011-05-23).
  \item The source code of the Sage samples in this script is delivered as
        Sage program files within the CrypTool-1 setup program.
        After installing CrypTool 1 \index{CrypTool 1} you find
        them within the subdirectory \verb#sage# within the CrypTool-1
        directory:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Index}
\printindex

\end{document}