### 2012-09-28 script/en/chapters/introduction.tex
Many small corrections and enhancements in the CT Script

% $Id$
% ............................................................................
%      V O R W O R T  und  E I N F Ü H R U N G (Zusammenspiel Skript-CT) 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% --------------------------------------------------------------------------
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Preface to the 11th Edition of the CrypTool Script}
\chapter*{Preface to the 11th Edition of the CrypTool Script}

Starting in the year 2000 this script became part of the CrypTool 1
(CT1)\index{CrypTool 1} package. The script's goal is to explain some mathematical
topics in more detail, but still in a way which is easy to understand.

Topics from mathematics and cryptography have been meaningfully split up and
for each topic an extra chapter has been written which can be read on its own.
This enables developers/authors to contribute independently of each other.
Naturally there are many more interesting topics in cryptography which could
be discussed in greater depth -- therefore this selection is only one of many
possible ways.

The later editorial work in TeX added cross linkages between different sections and footnotes describing where you can find the according functions within
the CrypTool programs. The appendices \ref{s:appendix-menu-overview-CT1} etc. contain a complete overview
of\index{CrypTool 1}\index{CrypTool 2}\index{JCrypTool}
the \hyperlink{menu-overview-CT1}{CrypTool 1 menu tree},
the \hyperlink{template-overview-CT2}{CrypTool 2 templates} and
the \hyperlink{function-overview-JCT}{JCrypTool (JCT) functions}.

The rapid spread of the Internet has lead to intensified research in the
technologies involved, especially within the area of cryptography where a good
deal of new knowledge has arisen.

%This edition of the CT script again updated some topics:
This edition completely updated the TeX sources of the document, and of course
the content of the script was amended, corrected, and updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the definitions of the strength of security functions
        (chap. \ref{cm_Section_Security_Definitions}),
  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{template-overview-CT2}{for CT2} and
        \hyperlink{function-overview-JCT}{for JCT},
  \item further Sage scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        Sage. Sage becomes more and more the standard open-source CAS system,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about RSA fixed points (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
%  \item the chapter about homomorphic encryption.
\end{itemize}

The script was delivered the first time with the CrypTool-1 package in
version 1.2.01. Since then it has been expanded and revised in almost
every new version of CrypTool 1.

I am deeply grateful to all the people helping with their impressive
commitment who have made this global project so successful.
Thanks also to the readers who sent us feedback.
% Especially I would like to acknowledge the English language proof-reading
% of this script version done by Richard Christensen and Lowell Montgomery.

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\
\\
% [1.5\baselineskip]
% \enlargethispage*{2\baselineskip}
% \nopagebreak
Bernhard Esslinger
\\
\\
% [\baselineskip]
Frankfurt (Germany), September 2012

% --------------------------------------------------------------------------
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Introduction -- How do the Script and the Program Play together?}
\chapter*{Introduction -- How do the Script and the Program Play together?}

\textbf{This script}

This document is delivered together with the open-source program
CrypTool 1\index{CrypTool 1}.

The articles in this script are largely self-contained and
can also be read independently of the CrypTool programs.

Chapters  \ref{Chapter_ModernCryptography} (Modern Cryptography) and 
\ref{Chapter_EllipticCurves} (Elliptic Curves) require a deeper knowledge
in mathematics, while the other chapters should be understandable with a 
school leaving certificate.

The \hyperlink{appendix-authors}{authors}
have attempted to describe cryptography for a broad 
audience -- without being mathematically incorrect. We believe that this
didactic pretension is the best way to promote the awareness for IT
security and the readiness to use standardized modern cryptography.
\par \vskip + 15pt

\noindent \textbf{The program CrypTool 1\index{CrypTool}}

CrypTool 1 (CT1) is an educational program with a comprehensive
online help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface.

CrypTool 1\index{CrypTool 1} and the successor versions
CrypTool 2 (CT2)\index{CrypTool 2} and JCrypTool (JCT)\index{JCrypTool}
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

\noindent \textbf{Acknowledgment}

At this point I'd like to thank explicitly the following people who
particularly contributed to CrypTool\index{CrypTool}. They applied their
very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Hr.\ Dominik Schadow
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann,
         Prof.\ Torben Weis and Prof.\ Arno Wacker.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt (Germany), July 2012

% Local Variables:
% TeX-master: "../script-en.tex"
% End: