### 2013-01-14 script/en/chapters/authors.tex

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{appendix-authors}{}
\section{Authors of the CrypTool Script}\index{CrypTool}
\label{s:appendix-authors}

This appendix lists the authors\index{Authors} of this document.\\
Please refer to the top of each individual chapter for their contribution.

\begin{description}

\item[Bernhard Esslinger,] \mbox{}\\
Initiator of the CrypTool project, main author of this script, head Cryptography Competence Center / Infrastructure IT Security Solutions at Deutsche Bank, and professor for IT security and cryptography at the University of Siegen.
E-mail: besslinger@web.de, esslinger@fb5.uni-siegen.de.

---------

\item[Matthias B\"uger,] \mbox{}\\ 
Contributor to chapter~\ref{Chapter_EllipticCurves} (``\nameref{Chapter_EllipticCurves}''),
research analyst at Deutsche Bank.

\item[Bartol Filipovic,] \mbox{}\\
Original author of the CrypTool elliptic curve
implementation and the corresponding chapter in this script.

\item[Martin Franz,] \mbox{}\\
Author of chapter~\ref{Chapter_HomomorphicCiphers}
(``\nameref{Chapter_HomomorphicCiphers}'').

\item[Henrik Koy,] \mbox{}\\
Main developer and co-ordinator of CT1 development
since version 1.3; script reviewer and \TeX\ guru; cryptographer 
and project leader IT at Deutsche Bank.

\item[Roger Oyono,] \mbox{}\\
Implementer of the CT1 factorization dialog and original author
of chapter~\ref{Chapter_ModernCryptography} (``\nameref{Chapter_ModernCryptography}'').

\item[J\"org Cornelius Schneider,] \mbox{}\\
Design and support of CrypTool; crypto enthusiast and IT architect and
senior project leader IT at Deutsche Bank.

\item[Christine St\"otzel,] \mbox{}\\
Master of Business and Computer Science at the University of Siegen.

---------

\item[Johannes Buchmann,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}'').
%Vice-president of the Technische Universit\"at Darmstadt, Germany and
%Professor at the Departments of Computer Science and of Mathematics.
Johannes Buchmann holds the Chair for Theoretical Computer Science (Cryptography
and Computer Algebra) at the department of Computer Science of the Technische
Universit\"at Darmstadt TUD).  He is also a Professor at the department of Mathematics, and vice-president of the university.

\item[Alexander May,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}''), 
Professor at the department of mathematics (chair for cryptology and IT Security) of the Ruhr-Universit\"at Bochum.

\item[Erik Dahmen,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}''). 
%Researcher at the Department of Computer Science of the TUD.
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

\item[Ulrich Vollmer,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}'').
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

---------

\item[Minh Van Nguyen,] \mbox{}\\
Sage developer and documentation quality reviewer.

\end{description}