### 2016-08-23 script/en/chapters/ellipticcurves.tex

% $Id$
% ...........................................................................
%                 E L L I P T I C  C U R V E S
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{bibunit}[babalpha] %% alpha: Chapter bibliography shows authors abbreviation

\newpage
\hypertarget{Chapter_EllipticCurves}{}

\chapter{Elliptic Curves}
\label{Chapter_EllipticCurves}
\index{elliptic curve}
(\hyperlink{author_Bartol-Filipovic}{Bartol Filipovic} /
 \hyperlink{author_Matthias-Bueger}{Matthias Bueger} /  % B\"uger
 \hyperlink{author_Bernhard-Esslinger}{Bernhard Esslinger} /
 \hyperlink{author_Roger-Oyono}{Roger Oyono},
 Apr 2000,
 Updates: Dec 2001, Jun 2002, Mar 2003, Nov 2009, Aug 2013, Apr 2016)

% -----------------------------------------------------------------------------
\section{Elliptic-curve cryptography -- a high-performance substitute for RSA?}
\index{performance}
\label{ECAlternative}

In many business sectors secure and efficient data transfer is essential.
In particular, the RSA algorithm is used in many applications. Although the
security of RSA is beyond doubt, the evolution in computing power has
caused a growth in the necessary key length.  Today, 1024-bit RSA keys are
standard, but the GISA\index{GISA} (German Information Security Agency)
recommends the usage of 2048-bit keys from 2006 on (compare
section~\ref{SecurityRSA}).  The fact that most chips on smart cards cannot
process keys extending ca. 2000 bit shows that there is a need for
alternatives in the area of asymmetric cryptography. Elliptic-curve
cryptography (ECC) is such an alternative. They are used widely on smartcards.

The efficiency of a cryptographic algorithm depends on the key length and
the calculation effort that is necessary to provide a prescribed level of
security. The major advantage of ECC compared to RSA is that it requires
much shorter key lengths.

If we assume that the computing power increases by Moore's law (i.~e.\ it
doubles every 18 months)\footnote{%
Empirical knowledge by Gordon Moore, co-founder of Intel, 1965}%
, then the evolution of the key lengths for secure communication will be as
in figure~\ref{RSAKeylength}, which was generated from table 1 (on page 32 in
\cite{LeVe2001}).\footnote{%
Further information about key length comparison by Arjen Lenstra und Eric
Verheul, plus more modern evaluations till 2015 can be found in the
interactive website \url{http://www.keylength.com}.
}

% -> Figure 1
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.7]{figures/RSAKeyLength-2}
\caption{Prognosis of the key lengths regarded to be safe for RSA and
  for elliptic curves} 
\label{RSAKeylength}
\end{center}
\end{figure}

In addition, a digital signature can be processed 10-times faster with ECC
than with RSA.  However, verification of a given signature is still more
efficient with RSA than with ECC. Refer to
figure~\ref{ThousandBitMultiplications} (source: Dr.~J.\ Merkle,
Elliptic-Curve Cryptography Workshop, 2001) for a comparison.  The reason is that
RSA public keys can be chosen relatively small as long as the secret key is
long enough.

% -> Figure 2
\begin{figure}[ht]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.7]{figures/ECCRSA}
\caption{Comparison between signing and verification time for RSA and elliptic curves} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}

Nevertheless, thin clients like smart cards usually have to store the (long)
secret key and have to process a digital signature rather than verify one.
Therefore, there is a clear advantage in using ECC in terms of efficiency.
\par
\smallskip
Nowadays, the major problem with ECC implementations is the lack of
standardization. There is only one way to implement RSA, but there are many
ways for ECC: One can work with different sets of numbers, different (elliptic)
curves --- described by parameters\footnote{%
See chapter \ref{ECC-Crypto}
} --- , and a variety of representations of the elements on the curve. Each
choice has its advantages and disadvantages, and one can certainly construct
the most efficient for each application. However, this causes problems in
interoperability. But if all ECC-tools should be able to communicate with each
other, they will have to support all different algorithms, which might put the
advantage of efficient computation and the need of less storage capacity to the
contrary.

\begin{sloppypar}
Therefore, international standardization organizations like IEEE (P1363),
ASC (ANSI X9.62, X9.63), ISO/IEC as well as major players like RSA labs or
Certicom have recently started standardization initiatives. While the IEEE
only describes the different implementations, the ASC has explicitly stated
10 elliptic curves and recommends their usage. The advantage of the ASC
approach is that one needs only a single byte to indicate which curve is
meant. However, it is not yet clear whether the ASC curves will become a de
facto standard.
\end{sloppypar}

Although there is no need to replace RSA in any application today\footnote{%
Current information about the security of the RSA algorithm can be found in 
chapters \ref{SecurityRSA} and \ref{Chapter_Dlog-FactoringDead}.},
one seriously should take the usage of ECC into consideration\footnote{%
Compare the technical guideline ``Cryptographic Methods: Recommendations and
Keylengths'' of GISA from February 15th, 2016.}.GISA\index{GISA}
More current discussions about the security of ECC can be found in
chapter \ref{Chapter_Dlog-FactoringDead}.

% -----------------------------------------------------------------------------
\section{Elliptic curves -- history}

Mathematicians have been researching elliptic curves for over 100 years.
Over the course of time, many lengthy and mathematically complex results
have been found and published which are connected to elliptic curves. 
A mathematician would say that elliptic curves (or the mathematics behind
them) are widely understood. This research was originally purely 
mathematical. That is to say, elliptic curves were investigated, for 
example, in the mathematical areas of number theory and algebraic 
geometry, which are generally highly abstract. Even in the recent
past, elliptic curves played an important role in pure mathematics.
In 1993 and 1994\footnote%
{
  1994 the gaps in the first proof have been closed by Wiles and Richard Taylor.
}
, Andrew Wiles\index{Wiles, Andrew} published 
mathematical works that triggered enthusiasm far beyond the specialist
audience. In these works, he proved a conjecture put forward in the 1960's.
To put it short, this conjecture was concerned with the connection 
between elliptic curves and what are called module forms. What is
particularly interesting for most people is that the works of Wiles
also proved the famous second theorem of Fermat\index{Fermat!last theorem}.
Mathematicians had spent centuries (Fermat\index{Fermat, Pierre} lived from
1601 to 1665) trying to find a
strict proof of this theorem. Understandably, therefore, Wiles' proof
got a good response. Fermat formulated his theorem as follows 
(written in the borders of a book from Diophantus):

\begin{quote} {\em
Cubum autem in duos cubos, aut quadratoquadratum in duos quadratoquadratos, et
generaliter nullam in infinitum ultra quadratum potestatem in duos ejusdem
nominis fas est dividere: cujus rei demonstrationem mirabilem sane detexi. Hanc
marginis exiguitas non caperet.
} \end{quote}

With a free translation, using the denotation of modern mathematics, this means:\\
No positive whole numbers $x, y$ and $z$ greater than zero exist such that
$x^n +y^n = z^n$ for $n>2$. I have found an amazing proof of this fact, but there is
too little space within the confines of this book to include it.

This is truly amazing: A statement that is relatively simple to understand (we
are referring to Fermat's second theorem here) could only be proved after such a
long period of time, although Fermat himself claimed to have found a proof.
What's more, the proof found by Wiles is extremely extensive (all of Wiles
publications connected with the proof made up a book in themselves). This should
therefore make it obvious that elliptic curves are generally based on highly
complex mathematics.

Anyway, that's enough about the role of elliptic curves in pure mathematics. 
In 1985 Neal Koblitz\index{Koblitz, Neal} and Victor Miller\index{Miller, Victor}
independently suggested using elliptic curves in cryptography. Elliptic 
curves have thus also found a concrete practical application. 
Another interesting area of application for elliptic curves is for
factorizing whole numbers \index{factorization} (the RSA cryptographic system
is based on the \index{complexity} difficulty/complexity of finding prime
factors of an extremely large number;  compare section \ref{SecurityRSA}.).
In this area, procedures based on elliptic curves have been investigated and
used since 1987 (compare section \ref{ECC-Factorization}).\\
There are also prime number tests\index{prime number!test} based on elliptic
curves.
% (s. Anmerkung U. Kuehn)

Elliptic curves are used differently in the various areas: Encryption
procedures based on elliptic curves are based on the difficulty of the problem
known as elliptic curve discrete logarithm\index{logarithm problem!discrete}.
The factorization of natural composite numbers $n$ uses the fact that a large
number of elliptic curves can be generated for $n$.%; however, these curves
%then are not groups for composite $n$.

% -----------------------------------------------------------------------------
\section{Elliptic curves -- mathematical basics}

This section provides information about \index{group} {\em groups} and
\index{field} {\em fields}.\footnote{%
  A didactically well prepared introduction into Elliptic Curves can be found
  in \cite{Schulz2015}.\index{RSA \& Co. at school}
}

% -----------------------------------------------------------------------------
\subsection{Groups}

Because the term {\em group} is used differently in everyday language than in
mathematics, we will, for reasons of completeness, begin by introducing the
essential statement of the formal definition of a group:
\begin{itemize}
   \item A group is a non-empty set $G$ on which an operation ``$\cdot$''.
	 The set $G$ is closed under this operation, which means that for
	 any two elements $a, b$ taken from $G$, performing the operation
	 on them gives an element in $G$, i.e. $ab=a\cdot b$ lies in $G$.
   \item For all elements $a, b$ and $c$ in $G$: $(ab)c = a(bc)$ (associative law).
   \item There exists an element $e$ in $G$ that behaves neutrally with respect
	 to the operation $\cdot$. That means that for all a in the set $G: ~ae = ea = a$.
   \item For each element $a$ in $G$ there exists a so-called
	 inverse\footnote{The inverse is uniquely determined because if $x,y\in G$
	 are each inverse to $a$, i.e. $ax=xa=e$ and $ay=ya=e$, then
         $x=xe=x(ay)=(xa)y=ey=y$.}
	 element $a^{-1}$ in $G$ such that: $aa^{-1} = a^{-1}a = e$.
\end{itemize}

If in addition it applies $ab = ba$ (commutative law) for all $a, b$ in $G$, then we call
the group an {\em Abelian} group.

Since we may define different operations on the same set, we distinguish
them by giving them different names (e.g.\ $+$ addition or $\cdot$ multiplication).

The simplest example of an (Abelian) group is the group of whole numbers under
the standard operation of addition. The set of whole numbers is denoted as
${\mathbb Z}$. ${\mathbb Z}$ has an infinite number of elements:
${\mathbb Z} = \{ \cdots, -4, -3, -2, -1, 0, 1, 2, 3, 4, \cdots\}$. For example, the
operation of $1+2$ lies in ${\mathbb Z}$, for $1+2 = 3$ and $3$ lies in
${\mathbb Z}$. The neutral element in the group ${\mathbb Z}$ is $0$. The
inverse element of $3$ is $-3$, for $3+(-3) = 0$.

For our purpose, so-called {\em finite} groups play an important role. This means
that there exists a set $\mathcal{M}$ with a fixed number of elements and an
operation $+$ such that the above conditions are fulfilled. One example is the
group ${\mathbb Z}_n = \{0, 1, 2, 3, \cdots, n-1\}$ of the remainders of the
division by $n \in {\mathbb N}$, and the operation is an addition mod $n$.
So, for example, $a$ and $b$ in ${\mathbb Z}_n$ are subject to the operation
$a+b \;{\rm mod~} n$.

\paragraph*{Cyclic groups}\index{group!cyclic}
Cyclic groups\footnote{Cyclic groups can be in general also endless like
the additive group of the integer numbers. We consider here only finite
cyclic groups.} are those groups $G'$ that possess an element $g$
from which the group operation can be used to generate all other
elements in the group. This means that for each element $a$ in
$G'$ there exists a positive whole number $i$ such that if $g$ is
subject to the operation $i$ times (i.e. ``$g\cdot i$''),
$g+g+\cdots+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of
the cyclic group --- each element in $G'$ can be generated using
$g$ and the operation.

\paragraph*{Group order}
Now to the order of an element of the group: Let $a$ be in $G$. The smallest
positive whole number $r$ for which $a$ subject to the operation with itself $r$
times is the neutral element of the group $G'$ (i.e.: $r \cdot a = a+a+\cdots+a =
e$ respectively $a^r = e$), is called the {\em order} of $a$.

The order of the group is the number of elements in the set $G$.

% -----------------------------------------------------------------------------
\subsection{Fields}

In mathematics, one is often interested in sets on which at least two (group)
operations are defined --- frequently called addition and multiplication.
Most prominent are so called fields.

A field is understood to be a set $K$ with two operations
(denoted as $+$ and $\cdot$) which fulfils the following conditions:
\begin{itemize}
   \item The set $K$ forms an Abelian group together with the operation $+$
(addition), where $0$ is the neutral element of the operation $+$.
   \item The set $K\setminus\{ 0\}$ also forms an Abelian group
together with the operation $\cdot$ (multiplication).
   \item For all elements $a, b$ and $c$ in $K$, we have $c\cdot (a+b) = c \cdot a + c
\cdot b$ and $(a+b) \cdot c = a \cdot c + b \cdot c$ (distributive law).
\end{itemize}

Fields may contain an infinite number of elements (e.g.\ the field of real numbers).
They are called {\em infinite} fields. In contrast we call a field
{\em finite}, if it contains only a finite number of elements
(e.g.\ ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$, where $p$
is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$).
\index{field!characteristic}
\paragraph*{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. Then the characteristic
of $K$ is said to be the order of $1$ with respect to the additive operation.
This means that the characteristic of $K$ is the smallest positive integer $n$
such that $$ \underbrace{1+1+\dots+1}_{\hbox{$n$ times}} =0 .
$$
If there is no such $n$, i.e. if $1+1+\dots+1\ne 0$ no matter how many $1$s we add, then we call $K$ a field
with characteristic $0$.

Thus, fields with characteristic $0$ are infinite since they contain the
(pairwise distinct) elements $1$, $1+1$, $1+1+1$, \dots. On the other hand,
fields with finite characteristic may by finite or infinite.

If the characteristic is finite, it has to be prime. This fact can easily be proved: Assume $n=pq$, $p,q<n$, is the characteristic of a field $K$. By definition of $n$, the elements $\bar p=\underbrace{1+1+\dots+1}_{\hbox{$p$ times}}$, $\bar q=\underbrace{1+1+\dots+1}_{\hbox{$q$ times}}$ of $K$ are not equal to $0$. Thus, there exist inverse elements $\bar p^{-1},\bar q^{-1}$ with respect to multiplication. It follows that $(\bar p\bar q)(\bar p^{-1}\bar q^{-1})=1$, which contradicts the fact that $\bar p\bar q=\bar n=\underbrace{1+1+\dots+1}_{\hbox{$n$ times}}=0$ and, hence, $\underbrace{(\bar p\bar q)}_{=0}(\bar p^{-1}\bar q^{-1})=0$.

\begin{remark}{:}\\
The field of real numbers has the characteristic $0$; the field ${\mathbb Z}_p$ has
the characteristic $p$. If $p$ is not prime, ${\mathbb Z}_p$ is not a field at all.
\end{remark}

The most simple field is ${\mathbb Z}_2=\{ 0,1\}$. It contains only two elements, the neutral elements with respect to addition and multiplication. In particular, we have $0+0=0$, $0+1=1+0=1$, $1+1=0$, $1\cdot 1=1$, $0\cdot 0=0\cdot 1=1\cdot 0=0$.

\index{field!finite}
\paragraph*{Finite Fields}
As mentioned above, each finite field has a characteristic $p\ne 0$, where $p$ is a prime. On the other hand, given a prime $p$ there is a field which has exactly $p$ elements, that is ${\mathbb Z}_p$.

However, the number of elements of a field need not be prime in general. For example, it is not hard to construct a field with $4$ elements\footnote{%
The set $K=\{0,1,a,b\}$ fitted with the operation defined in the tabular below is a field:\\
$
\begin{array}{|c||c|c|c|c|} 
\hline 
+ & 0 & 1 & a & b\\
\hline \hline
0 & 0 & 1 & a & b\\
\hline 
1 & 1 & 0 & b & a\\
\hline 
a & a & b & 0 & 1\\
\hline 
b & b & a & 1 & 0\\
\hline 
\end{array} \qquad {\rm ~und~} \qquad
\begin{array}{|c||c|c|c|c|} 
\hline 
\cdot & 0 & 1 & a & b\\
\hline \hline
0 & 0 & 0 & 0 & 0\\ 
\hline 
1 & 0 & 1 & a & b\\ 
\hline 
a & 0 & a & b & 1\\ 
\hline 
b & 0 & b & 1 & a\\
\hline 
\end{array} 
$\\
}.

One can show that the order of any field is a prime power (i.e. the power of a prime number). On the other hand, we can construct a field with $p^n$ elements for any given prime $p$ and positive integer $n$. Since two fields that have the same number of elements can not be distinguished\footnote{If $K,K'$ are fields with $k=p^n$ elements, then there is a one-to-one map $\varphi:K\to K'$, that respects the arithmetic of the field. Such a map is called an isomorphy. Isomorphic fields mathematically behave in the same way so that   it makes no sense to distinguish between them. For example, ${\mathbb Z}_2$ und $K'=\{ ZERO,ONE\}$ with zero-element $ZERO$ and one-element $ONE$ are isomorphic. We note that mathematical objects are only defined by their mathematical properties.}, we say that there is {\bf the field with $p^n$ elements} and denote it by $GF(p^n)$ or by $\mathbb{F}_p^{n}$ (used mostly in the English-speaking world). Here $GF$ stands for {\it Galois Field} to commemorate the French Mathematician Galois.

The fields $GF(p)$ of prime order play a prominent role. They are called prime fields and often also denoted by ${\mathbb Z}_p$.\footnote{%
  For prime fields additive as well as multiplicative group are cyclic. Furthermore,
  each field $GF(p^n)$ contains a subfield that is isomorphic to the prime field ${\mathbb Z}_p$.
}

% -----------------------------------------------------------------------------
\section{Elliptic curves in cryptography}\label{ECC-Crypto}

In cryptography elliptic curves are a useful tool.\footnote{%
  We write ``elliptic-curve cryptography'' with a hyphen, according to ``public-key cryptography''.
  Sadly, this isn't used consistently in the literature.
}

Such curves are described by some equation. A detailed analysis has shown that curves
of the form\footnote{%
  This curve is given by the zeros of a {\it polynomial}\index{polynomial}
  $F$ of degree three in three variables. In general, expressions of the form
  %%  $P=\sum_{i_1,\dots,i_n\in\mathbb{N}} a_{i_1\dots i_n} x_1^{i_1}\dots x_n^{i_n}$
  $P=\sum_{i_1,\dots,i_n} a_{i_1\dots i_n} x_1^{i_1}\dots x_n^{i_n}$
  with $i_1,\dots,i_n \in \mathbb{N}$
  with coefficients $a_{i_1\dots i_n}\in K$ are called polynomials in $n$ variables
  $x_1,\dots,x_n$ with underlying field $K$, if ${\rm deg\,} P:=\max\{i_1+\dots +i_n:
  a_{i_1\dots i_n}\ne 0\}$ is finite, i.e. the sum has only finitely many non-zero
  terms (monomials). The sum of the exponents of the variables of each term of the
  sum is at most $3$, at least one term of the sum has a single variable with $3$ as
  value of the according exponent.
}

\begin{equation}
 F(x_1,x_2,x_3)=-x_1^3+x_2^2x_3+a_1x_1x_2x_3-a_2x_1^2x_3+a_3x_2x_3^2-a_4x_1x_3^2-a_6x_3^3=0,
\label{eccbasisgleichung}
\end{equation}

are especially useful. The variables $x_1,x_2,x_3$ and parameters $a_1,\dots,a_4,a_6$ are elements of a given field $K$, which has certain properties that are make it useful from the cryptographic point of view. The underlying field $K$ might be the well known field of real numbers or some finite field (see last section).
In order to obtain a cure that is useful for cryptography, the parameters have to be chosen in a way that the following conditions hold
$$ \frac{\partial F}{\partial x_1}\ne 0, \quad \frac{\partial F}{\partial x_2}\ne 0, \quad
\frac{\partial F}{\partial x_3}\ne 0 .
$$
We identify points on the curve that can be derived from each over by multiplying each component with some scalar. This makes sense since $(x_1,x_2,x_3)$ solves (\ref{eccbasisgleichung}) if and only if $\alpha (x_1,x_2,x_3)$ ($\alpha\ne 0$) does. Formally, this means that we consider classes of equivalent points instead of single points, where points are called equivalent if one is the scalar multiple of the other one.
\\ If we put $x_3=0$ in the basic equation (\ref{eccbasisgleichung}), then this equation collapses to $-x_1^3=0$, leading to $x_1=0$. Thus, the equivalence class which includes the element $(0,1,0)$ is the only one that contains a point with $x_3=0$. For all points on the curve that are not equivalent to $(0,1,0)$, we may apply the following transformation
$$ K\times K\times (K\setminus\{0\})\ni (x_1,x_2,x_3) \mapsto (x,y):=\left( \frac{x_1}{x_3}, \frac{x_2}{x_3}\right) \in K\times K \, ,
$$
which reduces the number of variables to two instead of three. We note that the
basic equation (\ref{eccbasisgleichung}) $F(x_1,x_2,x_3)=0$  was chosen in a
way that this transformation leads to the famous so-called
Weierstrass-Equation\footnote{Karl Weierstrass\index{Weierstrass, Karl}, 31.10.1815$-$19.12.1897, German
mathematician, famous for his rigorous formal approach to mathematics.} holds
\begin{equation}
 y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6 \, .
\label{ell}
\end{equation}
Since all but one point (i.e. equivalence class) of the elliptic curve can be described using equation (\ref{ell}), this equation is often called the elliptic equation, and its solutions written as
$$ {\bf E} = \left\{(x,y)\in K\times K \, |\, y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6  \right\} \cup \{{\cal O} \}.
$$
Here, ${\cal O}$ represents the point $(0,1,0)$ that is loosely speaking mapped to infinity by the transformation (division by $x_3$) that reduces the three variables to two.

% -> Figure 3
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.60]{figures/elliptic-curve}
\caption{Example of an elliptic curve with the real numbers as underlying field} 
\label{ExampleEllipticCurve}
\end{center}
\vskip -10 pt
\end{figure}

In contrast to figure \ref{ExampleEllipticCurve} only finite fields $K=GF(p^n)$ are used in elliptic-curve cryptography. The reason is loosely speaking that in modern communication engineering data processing is always based on discrete data (simply because computers accept only discrete data).

For practical reasons, it turned out to be useful to take either $GF(p)$ with a large prime $p$ or $GF(2^n)$ with a (large) positive integer $n$. Using $GF(p)$ has the advantage of providing a relatively simple arithmetic; on the other hand $GF(2^n)$ allows a binary representation of each element that supports the way computers work. Other fields like, for example, $GF(7^n)$ do not have any of these advantages and are, thus, not considered, although there is no mathematical reason why they should not.

A coordinate transformation can result in a simpler version\footnote{Such a coordinate transformation is combination of a rotation and a dilatation of the coordinate system without changing the elliptic curve itself.} of the Weierstrass equation\index{Weierstrass, Karl}. Depending whether $p>3$, different transformations are used, and we obtain
\begin{itemize}
\item in case of $GF(p)$, $p>3$, the elliptic curve equation of the form
\begin{equation}
 y^2 = x^3 + ax + b
\label{ellp}
\end{equation}
with $4a^3+27b^2\ne 0$
\item in case of $GF(2^n)$ the elliptic curve equation of the form 
\begin{equation}
 y^2+xy = x^3 + ax^2 + b
\label{ell2}
\end{equation}
with $b\ne 0$\footnote{The form (\ref{ellp}) is called the standard form of the Weierstrass-equation\index{Weierstrass, Karl}. If the characteristic of the field is $2$ or $3$, we obtain $4=0$ respectively $27=0$, which means that the condition on parameters $a,b$ collapse. Loosely speaking, this is the reason why the transformation to the standard form does not work in these cases.}.
\end{itemize}
This conditions on the parameters $a,b$ ensure that the elliptic equation can be used in the context of cryptography\footnote{Formally we call such curves non singular.}.

Let $|E|$ denote the number of elements of an elliptic curve $E$ given an underlying field $GF(k)$ (for practical reasons either $k=p$ with $p$ prim or $k=2^n$). Then Hasse's theorem\cite{Silverman2009} yields $| \, |E| - k-1\,| \le 2\cdot \sqrt{k}$. This inequality is equivalent to $k+1 - 2\sqrt{k} < |E| < k+1+2\sqrt{k}$. In particular, this means that the number of elements of an elliptic curve is approximately $k$ (for large $k$).

% -----------------------------------------------------------------------------
\section{Operating on the elliptic curve}

In order to work with elliptic curves in practice, we define an operation (often written in an additive way $+$) on the set of points on the curve. If we have a curve over the field $GF(p)$, we define the commutative operation $+$ by
\begin{enumerate}
\item $P+{\cal O}={\cal O}+P=P$ for all $P\in E$,
\item for $P=(x,y)$ and $Q=(x,-y)$ we set $P+Q={\cal O}$,
\item for $P_1=(x_1,x_2),P_2=(x_2,y_2)\in E$ with $P_1,P_2\ne {\cal O}$ and $(x_2,y_2)\ne (x_1,-y_1)$ we set $P_3:=P_1+P_2$, $P_3=(x_3,y_3)$ defined by
$$ x_3:=-x_1-x_2+\lambda^2 \, , \qquad y_3:=-y_1+\lambda (x_1-x_3)
$$
with the auxiliary quotient
$$ \lambda:=\left\{ \begin{array}{cl} \frac{y_1-y_2}{x_1-x_2} & {\rm if~} P_1\ne P_2,\\
                                     \frac{3x_1^2+a}{2y_1} & {\rm if~} P_1=P_2. \end{array} \right.
$$
\end{enumerate}
In particular, we obtain $-P=(x,-y)$ for $P=(x,y)\in E$.

If we deal with a curve over the field $GF(2^n)$, we define the operation $+$ in an analogous way by
\begin{enumerate}
\item $P+{\cal O}={\cal O}+P=P$ for all $P\in E$,
\item for $P=(x,y)$ and $Q=(x,x+y)$ we set $P+Q={\cal O}$,
\item for $P_1=(x_1,x_2),P_2=(x_2,y_2)\in E$ with $P_1,P_2\ne {\cal O}$ and $(x_2,y_2)\ne (x_1,x_1+y_1)$ we set $P_3:=P_1+P_2$, $P_3=(x_3,y_3)$ defined by
$$ x_3:=-x_1+x_2+\lambda+\lambda^2+a \, , \qquad y_3:=y_1+x_3+\lambda (x_1+x_3)
$$
with auxiliary quotient
$$ \lambda:=\left\{ \begin{array}{cl} \frac{y_1+y_2}{x_1+x_2} & {\rm if~} P_1\ne P_2,\\
                                   x_1+\frac{y_1}{x_1} & {\rm if~} P_1=P_2. \end{array}\right.
$$
\end{enumerate}
In particular, we obtain $-P=(x,-y)$ for $P=(x,y)\in E$.

(Note that $-(-P)=(x,x+(x+y))=(x,2x+y)=(x,y)$, since the underlying field
has characteristic $2$.)\footnote{%
An animation of the addition of points
on elliptic curves can be found on the Certicom\index{Certicom} homepage\\
\url{https://www.certicom.com/ecc-tutorial} (Date of last update is unknown).\\
See also the web link about the \hyperlink{ec:Web-Link:Java_Laubrock}
{Java-Tutorial} at the end of this chapter.
}

One can verify that $+$ defines a group operation on the set $E\cap\{\cal O\}$.
In particular this means that the sum of two points is again a point on the
elliptic curve. How his operation works is geometrically visualized in the
following section.

% \newpage
\begin{figure}[htbp]
\subsection*{How to add points on an elliptic curve}
The following figures show how points on an elliptic curve over the field of real numbers are summed up using affine coordinates.
We note that the point infinity ${\cal O}$ cannot be shown in the affine plane.  
\begin{center}
\includegraphics[scale=1.08]{figures/ec-mult2}
\caption{Doubling of a point} 
\vspace{\floatsep}
\vskip +20 pt
\includegraphics[scale=0.65]{figures/ec-add}
\caption{Summing up two different points over the real number field} % \footnotemark }
\end{center}
\end{figure}
\enlargethispage{+20pt}
\newpage

% -----------------------------------------------------------------------------
\section{Security of elliptic-curve cryptography: The ECDLP}

As mentioned above in section \ref{ECC-Crypto}, we only consider elliptic curves over the finite\footnote{Discrete in contrast to continuous.} fields $GF(2^n)$ or $GF(p)$ (for a large prime $p$). This means that all parameters that describe the curve are taken from this underlying field. If $E$ is an elliptic curve over such a field and $P$ is a point on the curve $E$, then we can derive for all positive integers $m$
$$ mP := \underbrace{P+P+\dots+P}_{\hbox{$m$ times}} \, .
$$
Looking on this operation from the cryptographic point of view, it turns out to be very interesting by the following reason: On the one hand one needs only $\log m$ operations to calculate $mP$ --- one simply has to calculate $P$, $2P$, $2^2P$, $2^3P$, \dots, write $m$ in a binary form and finally add all these multiples $2^kP$ of $P$ with respect to the binary representation of $m$ --- on the other hand it seems to be very hard to find $m$ given $P$ and $Q=mP$ on $E$. Of course, we may simply calculate $P,2P,3P,4P,5P,\dots$ and compare each of them with $Q$. But this will take as much as $m$ operations.

Yet there is no algorithm known that efficiently derives $m$ given $P$ and $G$. The best algorithms known so far need about $\sqrt{q}$ operations where $q$ is the (largest) prime factor of $p-1$, in case the underlying field is $GF(p)$; here $m$ should be between $1$ and $q$ liegen so that one needs at most $\log q$ operations to calculate $mP$. However, the quotient $\frac{\sqrt{q}}{\log q}$ tends to $+\infty$ very fast for large $q$.

If we choose the parameters sufficiently large (for example, let $p$ be prime and at least $160$ bits long), an computer will easily be able to calculate $mP$ (in less than a second). The {\it inverse problem} however, to derive $m$ from $mP$ and $P$, can (still) not be solved in reasonable time.

This problem is known as the ``Elliptic Curve Discrete Logarithm Problem'' (for short ECDLP\index{ECDLP}).

\vskip +5 pt

In elliptic-curve cryptography we formally look at points on the elliptic curve as elements of a group with point addition $+$ as operation. Furthermore, we use only elliptic curves that have a sufficiently large number of points. However, in special cases curves may be weak and not useful due to other reasons. For such special cases the ECDLP can be much easier to solve than in the general case. This means that one has to look carefully at the parameters when choosing an elliptic curve for cryptographic applications.

Not useful for cryptography are {\em a-normal} (that are curves over ${\mathbb Z}_p$,
for which the set ${\bf E}$ consists of exactly $p$ elements) and {\em supersingular} curves (that are curves, for which the ECDLP can be reduced to the ``normal'' discrete logarithms in another, smaller finite field). This means that there are cryptographically useful and non-useful elliptic curves. Given the parameters $a$ and $b$, it is possible to determine whether a curve is useful or not. In many publications one can find parameters that turned out to be useful for cryptography. The open (scientific) discussion guarantees that these results take into account latest research.
\vskip +5 pt

Given a secure curve, the time that is needed to solve the ECDLP is strongly correlated with parameter $p$ in case $GF(p)$ respectively $n$ in case of $GF(2^n)$. The larger these parameters become, the more time an attacker needs to solve the ECDLP --- at least with the best algorithms known so far. Experts recommend bit-lengths of $200$ for $p$ for secure curves. A comparison with RSA modulus length shows why elliptic curves are so interesting for applications. We note that the computation effort for signing and encryption is closely related to the bit-length of the parameters. In addition the initiation process, i.e. the generation of the private-public-key-pair, becomes more complicated the larger $p$ is. Thus, one looks for the smallest parameters that still come along with the security required. It is remarkable that a length of $200$ bits for $p$ is sufficient to construct a {\em good} elliptic curve that is as secure as RSA with a $1024$ bit \index{RSA!modulus} RSA modulus (as far as we know today). For short, the reason for this advantage of ECC lies in the fact that the best algorithms known for solving the ECDLP need exponential time while the best algorithms for factorizing are sub-exponential (number field sieve, quadratic sieve or factorizing with elliptic curves). Hence, the parameters for a cryptosystem that is based on the problem of {\em factorizing large integers} have to be larger than the parameters for a system based on ECDLP.

% -----------------------------------------------------------------------------
\section{Encryption and signing with elliptic curves}

\begin{sloppypar}
  The {\em elliptic curve discrete logarithm
    problem}\index{problem of discrete logarithm} (ECDLP) \index{ECDLP} is the basis for elliptic-curve cryptography. Based on this problem, there are different signature schemes. In order to apply one of these, we need:
\end{sloppypar}
\begin{itemize}
    \item An elliptic curve {\bf E} with an underlying field $GF(p^n)$.
    \item A prime $q\ne p$ and a point $G$ on the elliptic curve ${\bf E}$ with order $q$. This means that $qG={\cal O}$ and $rG\ne {\cal O}$ for all $r\in \{1,2,\dots,q-1\}$. Thus $q$ is a factor of the group order (i.e. the number of elements) $\#{\bf E}$ of $E$. Since $q$ is prime, $G$ generates a cyclic sub-group of ${\bf E}$ of order $q$.
\end{itemize}
The parameters mentioned are often called \index{domain parameter}
{\em Domain} parameter. They describe the elliptic curve ${\bf E}$ and 
the cyclic sub-group of ${\bf E}$ on which the signature scheme is based.

\par
%\smallskip
%{\bf Encryption:}
\subsection{Encryption}

Using elliptic curves one can construct a key exchange protocol based on the Diffie-Hellman protocol \index{Diffie-Hellman} (see chapter \ref{DH-KeyExch}). The key exchanged can be used for a subsequent symmetric encryption. We note that in contrast to RSA there is no pair of private and public key that can be used for encryption and decryption!

In the notation of elliptic curves, the Diffie-Hellman protocol reads as follows: First both partners (A und B) agree on a group $G$ and an integer $q$. Then they choose $r_A,r_B\in\{1,2,\dots,q-1\}$ at random, derive the points $R_A=r_AG$, $R_B=r_BG$ on the elliptic curve and exchange them (using an insecure channel). After that A easily obtains $R=r_AR_B$; B gets the same point ($R=r_Ar_B G$) by calculating $r_BR_A=r_Br_AG=r_Ar_BG=R$. We note that $R_A,R_B$ are easy to derive as long as $r_A$ respectively $r_B$ are known $G$. However, the inverse operation, to get $R_A$ respectively $R_B$ from $r_A$ respectively $r_B$ is hard.
\\ Using the best algorithms known so far, it is impossible for any attacker to obtain $R$ without knowing either $r_A$ or $r_B$ --- otherwise he would have to solve the ECDLP.

In order to prohibit a ``Man-in-the-middle"{}~attack, one may sign the values $G,q,R_A,R_B$ as described in chapter \ref{Impersonalisierungsattacke}.

\par
%\smallskip 
%{\bf Signing:}
\subsection{Signing}

Using the \index{DSA} DSA signature scheme, one can proceed as follows: The signing party chooses a (non-trivial) number $s\in{\mathbb Z}_q$, which will be the private key, and publishes $q$, $G$ and $R=sG$. We note that $s$ cannot be obtained from $G$ and $R$ are not sufficient --- a fact on which the security of the signature scheme is based.

Given the message $m$, which should be signed, one first constructs a digital finger print using a hash-algorithm $h$ such that $h(m)$ has its values in $\{0,1,2,\dots, q-1\}$. Thus, $h(m)$ can be considered as an Element of ${\mathbb Z}_q$. Then the signing party chooses a random number $r\in{\mathbb Z}_q$ and derives $R=(r_1,r_2)=rG$. We note that the first component $r_1$ of $R$ is an element of $GF(p^n)$. This component will then be projected onto ${\mathbb Z}_q$, i.e. in case of $n=1$ it is interpreted as the remainder of an element of $\{0,1,\dots,p-1\}$ divided by $q$. This projection of $r_1$ onto ${\mathbb Z}_q$ is denoted by $\bar r_1$. Then one determines $x\in {\mathbb Z}_q$ such that
$$ rx-s\bar r_1-h(m)=0 .
$$
The triple $(m,r_1,x)$ is then published as the digital signature of message $m$.

\par
%\smallskip 
%{\bf Signature Verification:}
\subsection{Signature verification}

In order to verify a signature, one has to build $u_1=h(m)/x$, $u_2=\bar r_1/x$ (in ${\mathbb Z}_q$ and derive
$$ V=u_1G+u_2Q .
$$
Since we have $Q=sG$, the point $V=(v_1,v_2)$ satisfies $v_1=u_1+u_2s$. We note that this operations take place in the field $GF(p^n)$. The projection of $GF(p^n)$ on ${\mathbb Z}_q$ mentioned above should be chosen in such a way that $\bar v_1=u_1+u_2s$ is an element of ${\mathbb Z}_q$. Then it follows that
$$ \bar v_1=u_1+u_2s=h(m)/x+\bar r_1 s/x=(h(m)+\bar r_1s)/x=rx/x=r .
$$
Since $R=rG$, we obtain $\bar v_1=\bar r_1$, i.e. $R$ and $V$ coincide modulo the projection onto ${\mathbb Z}_q$.

% -----------------------------------------------------------------------------
\hypertarget{faktell}{}
\section{Factorization using elliptic curves} \label{ECC-Factorization}

There are factorization%
\footnote{John M. Pollard \index{Pollard, John M.} was involved
in the development of many different factorization algorithms; also at
factorization with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contributions in mathematics''.\\
In 1987 H.W. Lenstra published an often used factorization algorithm,
based on elliptic curves (see \cite{Lenstra1987}).
}
algorithms based on elliptic curves%
\footnote{The biggest compound numbers currently 
factorized\index{factorization!factoring records} with elliptic curves have
about 80 decimal digits:\\
\url{https://members.loria.fr/PZimmermann/records/top50.html}.\\
See also the web link about the \hyperlink{Lenstra2}{ECMNET project}\index{ECMNET} at the end of this chapter.
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

Factorization using elliptic curves thus principally works as follows: 
Random curves over ${\mathbb Z}_n$ are selected, as well as random points
(that lie on this curve) and add them; you thus obtain points that also
lie on the curve or find a factor of $n$. 
Factorization algorithms based on elliptic curves
therefore work probabilistically. The opportunity of defining large number of
elliptic curves over ${\mathbb Z}_n$ allows you to increase the probability of
finding two points which you can add to obtain a factor of $n$. These procedures
are therefore highly suitable for parallelization.

% -----------------------------------------------------------------------------
\newpage
% \section{Implementing elliptic curves for educational purposes, pat\-ent aspects}
\section{Implementing elliptic curves for educational purposes}
\label{ec:Implementing-for-Education}

There are not many free programms offering ECC under a graphical user interface.
The following subsections explain which according functionality is available in
CrypTool and in SageMath.

% -----------------------------------------------------------------------------
\subsection{CrypTool}\index{CrypTool}

CT1 offers elliptic curves for the digital signature function%
\footnote{%
The dialog box, which appears in CT1\index{CT1} after clicking the menu
{\bf Digital Signatures/PKI \textbackslash{} Sign Message},
offers the EC methods ECSP-DSA and ECSP-NR.

} and for ECC-AES hybrid encryption%
\footnote{%
Within CT1\index{CT1} you can find this technique using the menu
path {\bf Crypt \textbackslash{} Hybrid}.

}.

It implements the basic algorithms for group operations, for generating elliptic
curves, for importing and exporting parameters for elliptic curves over finite
fields with $p$ elements ($p$ prime). The algorithms have been implemented in
ANSI C and comply with draft no. 8 of the IEEE P1363 work group {\em Standard
Specifications for Public Key Cryptography}

{\url{http://grouper.ieee.org/groups/1363}}.

The procedure implements the cryptographic primitives for generating and
verifying signatures for the variations of Nyberg-Rueppel signatures and
\index{DSA} DSA signatures based on elliptic curves.

Step-by-step point addition on elliptic curves is visualized in CT1\index{CT1}
and in JCT\index{JCT}.%
\footnote{%
  CT1: menu {\bf Digital Signatures/PKI \textbackslash{} Signature Demonstration
  (Signature Generation)},\\
  JCT (default perspective): menu {\bf Visuals \textbackslash{} Elliptic Curve
  Calculations}.

}

% -----------------------------------------------------------------------------
\subsection{SageMath}
\label{ec:Sage_Massierer}
\index{SageMath}
\index{SageMath!code examples}

In SageMath elliptic curves are very well described at:%
\footnote{%
According SageMath samples can be found e.g. also in the "Elliptic Curve Cryptography
(ECC) Tutorial"\\
\url{http://www.williamstein.org/simuw06/notes/notes/node12.html}

}

\begin{sloppypar}
\begin{itemize}
    \item \url{http://doc.sagemath.org/html/en/constructions/elliptic_curves.html}
    \item \url{http://doc.sagemath.org/html/en/reference/plane_curves/index.html#elliptic-curves}
\end{itemize}
\end{sloppypar}

\noindent Additionally there is an exhaustive, interactive
\hyperlink{ec:Web-Link:Sage_Massierer}{ECC tutorial} by Maike Massierer.
This interactive introduction to Elliptic-Curve Cryptography is built up as
a SageMath notebook.

\noindent SageMath notebooks are running after a logon within a browser%
\footnote{%
If you installed SageMath on your own (Unix) server, you first have to
enter at the command line the command \verb#notebook()#.

}${}^,$\footnote{%
The \hyperlink{ec:Web-Link:Sage_Massierer}{ECC notebook} of Massierer
needs the KASH3 library: Therefore e.g.\ with SageMath 4.2.1%TODO-Version
the package ``kash3-2008-07-31.spkg'' has to be installed (command \verb#sage -i#).
}.

%--------------
The \hyperlink{ec:Web-Link:Sage_Massierer}{ECC notebook}
\index{elliptic curve!ECC notebook}\index{Massierer, Maike}
created in 2008 by Massierer\footnote{%
Instructions to use an interactive SageMath notebook\index{SageMath!instructions
interactive notebook}:
Update for the new SageMathCloud xxxxxxxxxxxxx\\
% TODOTODO

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Some SageMath servers are publicly available
  %like \url{http://sage.mathematik.uni-siegen.de:8000} or
  %\url{http://www.sagenb.org/}
  and offer running samples as ``Published Worksheets'',
  which you can run and download without login on. These worksheets are listed
  if you click on ``Published'' in the above right corner.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Worksheets using the  \verb#interact# command currently need some additional
  todos for a user to work correctly: sign-in, make a copy and execute all
  commands again.\\
%  This works as follows (described for the sagenb server and for the ECC tutorial):\\
% 
%  TODO --> SageMathCloud xxxxxxxxxxx
%  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
%  Sign-up for a SageMath notebook account at \url{http://sagenb.org/register}
%  and sign in at \url{http://sagenb.org/}.

%  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
%  Open the worksheet \url{http://sagenb.org/home/pub/1126/}.
%  --> Now
%  https://cloud.sagemath.com/projects/19575ea0-317e-402b-be57-368d04c113db/files/pub/1101-1201/1126.sagews
%  This contains the table of contents of the interactive ECC notebook.
%  From here you can navigate via a click to the different chapters of the document.

%  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
%  In the very top left corner, click \verb#Edit a copy# in order to create
%  your own copy of the worksheet.

%  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
%  Sometimes it's necessary at the beginning to re-evaluate the worksheet.
%  Click in the left upper corner on \verb#Action -> Evaluate all#.

%  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
%  Some of the applications still do not always work after opening a worksheet.
%  Instead of nice output, they show lots of (blue) error messages.
%  This normally can be solved quickly by clicking the gray  ``\%hide'' string:
%  Then you get the code behind the graphics. Simply generate the graphics
%  again with Shift-Enter.\\
%  Even after doing this, the graphics code does not always disappear. Instead,
%  it sometimes turns gray. Should this happen, click on the gray text,
%  then click somewhere outside of the text box. The code will then
%  disappear and leave you with a nice layout of the worksheet.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Some of the ECC tutorial's content uses a special math fonts that are
  not installed by default with most browsers. When you notice that
  formulas are not displayed correctly or even get an error message
  about missing fonts from your browser, you need to install the jsMath
  fonts for a better layout.\\
  See \url{http://www.math.union.edu/~dpvc/jsMath/}.\\
  After installing these fonts you can see the jsMath symbol at the lower border
  of your browser. If you click this symbol you can find the download page for
  the TIFF fonts. This font installation has to be done at every PC.
%
%  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
%  According to the SageMath-support newsgroup there is work underway
%  to create a system for using \verb#@interact# completely outside of the
%  SageMath notebook (JS code within a static html pages).

}
consists of 8 parts (title page with contents plus 7 chapters)
and aims to let even beginners understand what elliptic curves are:
\begin{enumerate}
   \setcounter{enumi}{-1}
   \item ECC Notebook (title page and contents)
   \item Introduction and Overview
   \item Motivation for the use of Elliptic Curves in Cryptography
   \item Elliptic Curves in Cryptography
   \item Cryptographic Protocols in ECC
   \item Domain Parameter Generation for ECC Systems
   \item Conclusion and Further Topics
   \item References
\end{enumerate}

% -----------------------------------------------------------------------------
\newpage
\section{Patent aspects}\index{patent}

If the field $GF(2^n)$ is used instead of the prime field $GF(p)$, one has to make substantial changes in the implementation. The advantage of $GF(2^n)$ lies in the fact that calculations in $GF(2^n)$ can be implemented very efficiently using the binary representation. In particular, divisions are much easier to process compared to $GF(p)$ (this is particularly important in the signature scheme mentioned above where a division is needed for processing a signature as well as for the verification).

In order to achieve maximal gain in efficiency, one may choose a field that allows special basis like polynomial basis (useful for software implementations) or normal basis (best for hardware implementations). For special $n$ (like, for example, $n=163,179,181$) one may even combine both advantages. However, they are still non-standard.

Sometimes only the first component and one additional bit is used as representation of a point on the elliptic curve instead of the full two components. Since the first component together with the additional bit is sufficient to derive the full point, this representation minimizes the memory capacity needed. In particular, for normal basis this point compression can be implemented efficiently. In addition, the cryptographic protocols themselves become more effective. A disadvantage is, however, that {\it point compression} can be used for about half of all elliptic curves only and is protected under
US patent (US Patent 6141420, Certicon), causing additional costs.
In the general case $GF(p^n)$ (and also in case $n=1$) often so called affine or projective co-ordinates are used. Depending on the application, these co-ordinates may result in a gain in efficiency as well.

A comprehensive description of all implementations and their advantages and disadvantages would go far beyond the scope of this paper. We only want to state that there is a variety of possible implementations for elliptic-curve cryptography, much more than for RSA. Therefore, there are serious efforts to reduce this large to a small number of standard implementations. Some standardization committees even try to reduce the complexity by focusing on a small number of (prescribed) curves (ASC approach).

It is still not clear whether these standardization initiatives will be
successful or not. However, without agreed standards, ECC is not likely to
become a real alternative for RSA. The committees might be forced to act fast
if there was a break-through in factorization.

Current informationen about the patents situation can be found here\footnote{%
\url{https://en.wikipedia.org/wiki/Elliptic_curve_cryptography}\\
\url{https://en.wikipedia.org/wiki/ECC_patents}\\
\url{https://cr.yp.to/ecdh/patents.html} Bernstein, Daniel J. (2006-05-23):
``Irrelevant patents on elliptic-curve cryptography''. Retrieved 2016-07-14.
}.

% -----------------------------------------------------------------------------
\section{Elliptic curves in use}

% Einsatz in Protokollen wie TLS, Chats, ... ergänzen !!  xxxxxxxxxxxxxxxxxxxxxxxxx
Today elliptic-curve cryptography is already broadly in use. A prominent example is the
information network Bonn-Berlin\footnote{%
The Informationsverbund Bonn-Berlin (IVBB) connects governmental institutions
in the old and new German capital.
}\index{IVBB}, used for the exchange
of strictly confidential documents between different German federal governmental
institutions in Berlin and Bonn. With the help of ECC a high security solution
could be realized. Interoperability, however, played only a minor role.

In Austria ECC has been massively launched: A bank card with digital signature function.

Both examples show typical applications for elliptic-curve cryptography: For high security solutions and for implementations on smartcards in which the key length is crucial (because of lack of physical memory available).

%------------------------------------------------------------------------------
\putbib[../de/references]
\addcontentsline{toc}{section}{Bibliography}
\end{bibunit}

\noindent All links have been confirmed at July 14, 2016.

%--------------------------------------------------------------------
\chapter*{Web links}
\addcontentsline{toc}{section}{Web links}

\begin{enumerate}

   \hypertarget{ec:Web-Link:Sage_Massierer}{} 
   \item Interactive introduction to elliptic curves and elliptic curve
        cryptography with SageMath\index{SageMath} 
        by Maike Massierer and the CrypTool team,\\
        ``ECC Tutorial as SageMath notebook'', Version 1.3, January 2011\\
        \url{http://web.maths.unsw.edu.au/~maikemassierer/ecc-notebook/}

   \hypertarget{ec:Web-Link:Certicom_Tutorial}{} 
   \item Certicom Online Tutorial\index{Certicom},\\
        \url{https://www.certicom.com/index.php/ecc-tutorial}

   \hypertarget{ec:Web-Link:RSALabs}{}
   \item Overview of Elliptic Curve Cryptosystems,\\
        Revised June 27, 1997.~
        M.J.B. Robshaw and Yiqun Lisa Yin.\\
        RSA Laboratories\index{RSA Laboratories},
        \begin{sloppypar}
        \url{http://www.emc.com/emc-plus/rsa-labs/historical/overview-elliptic-curve-cryptosystems.htm}
        \end{sloppypar}

   \hypertarget{ec:Web-Link:Java_Laubrock}{} 
   \item Tutorial with Java applets -- Crypto methods based on elliptic curves,\\
        thesis by Thomas Laubrock, 1999 (German only),\\
        \url{http://www.warendorf-freckenhorst.de/elliptische-kurven/frame.html}

   \item Working group IEEE P1363,\\
        \url{http://grouper.ieee.org/groups/1363}

   \item \hypertarget{Lenstra2}{}
        An informative web page about factorization with elliptic curves,\\
        It contains literature related to the topic factorization with 
	elliptic curves as well as links to other web pages.\\
        \url{https://members.loria.fr/PZimmermann/records/ecmnet.html}

   \item BSI TR-02102-1,\\
	Technical Guideline for Cryptographic Methods: Recommendations and
        Keylengths by GISA\index{GISA} (German Information Security Agency),\\
	February 15th, 2016. (German only)
        % BSI -- Technische Richtlinie
        % Bezeichnung:  Kryptographische Verfahren: Empfehlungen und Schlüssellängen
        % Kürzel:       BSI TR-02102-1
        % Version:      2016-01
        % Stand:        15. Februar 2016
        \begin{sloppypar}
        \url{https://www.bsi.bund.de/EN/Topics/ElectrIDDocuments/TRandSecurProfiles/technicalGuidelinesSeite.html}\\
        \url{https://www.bsi.bund.de/SharedDocs/Downloads/DE/BSI/Publikationen/TechnischeRichtlinien/TR02102/BSI-TR-02102.pdf?__blob=publicationFile&v=2}
        \end{sloppypar}

\end{enumerate}

\noindent All links have been confirmed at July 14, 2016.

% Local Variables:
% TeX-master: "../script-en.tex"
% End: